### 2015-12-22 Master.Rnw
Master file for 1. round review version of the paper

\documentclass[12pt,english]{article}

\usepackage{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{url}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[authoryear]{natbib}
\usepackage[unicode=true]{hyperref}
\usepackage[format=hang, justification=raggedright]{subfig}
\usepackage{array}
\usepackage{multirow}
%\usepackage{authblk}
\usepackage{lscape}
\usepackage{float}
\usepackage{longtable}
\usepackage[noae]{Sweave}
\usepackage{setspace}
\doublespacing
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=3cm,rmargin=3cm}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Income and Food Price Transmission\footnote{Revised version of ``Poverty and Price Transmission''. Submitted to \emph{Agricultural Economics} for review}}

\author{Christian Elleby}
\date{2015}

\maketitle
\begin{abstract}
The paper analyzes how international-to-domestic food price transmission
depends on a country\textquoteright s income or expenditure level.
According to conventional wisdom, international food price spikes
have the largest poverty impact on low income countries. However,
the paper presents empirical evidence of an inverse-U shaped relationship
between the degree of international agricultural price changes passed
on to domestic consumers and the country\textquoteright s per capita
income or expenditure level. This implies that retail food prices
in middle income countries respond the strongest to international
commodity price spikes, whereas those in low income countries respond
less, all else equal. The specification of the empirical model is
based on a simple theoretical model of a country\textquoteright s
aggregate food price level. Its predictions are tested on monthly
price index data from 140 countries covering the six year period 2005-11
and results are found to be generally significant and robust. Around
one third to one half of the countries in the sample are to the left
of the turning point, on the increasing part of the parabola, in most
specifications.\end{abstract}

\SweaveInput{intro.Rnw}

\SweaveInput{litrev.Rnw}

\SweaveInput{model.Rnw}

\SweaveInput{analysis.Rnw}

\SweaveInput{conclusion.Rnw}

\bibliographystyle{chicago}
\bibliography{poverty_price_transmission}

\newpage

\appendix

\SweaveInput{figures.Rnw}

\newpage

\SweaveInput{tables.Rnw}

\end{document}