### 2015-12-22 conclusion.Rnw
Concl. sec of 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\section{Conclusion\label{sec:conclusion}}
This paper contributes to the literature on aggregate food price transmission with three findings. First, domestic food prices in middle income countries react the strongest to international agricultural commodity price changes once we control for differences in consumption and trade dependency patterns between high and low income countries. Secondly, domestic price increases were the largest in middle income countries during the two recent international price spikes in 2007-08 and 2010-11, all else equal. Thirdly, there is a simple economic logic behind this inverse-U shaped income-price transmission relationship. These findings are worrying due to two other stylized facts. The first is that middle income countries are now home to around three quarters of the world’s poor. The second is that poor households spend a large portion of their income on food and are therefore vulnerable to sudden increases in the price of food goods. There is, however, also a positive side to this finding. If, as some authors claim, middle income countries, in principle, have the domestic financial resources to eradicate extreme poverty \citep{Sumner2013}, then they should also possess the necessary resources to implement social safety nets which shield purchasing power of the poor from occasional price spikes on the international markets.

The analysis has at least two major policy implications. First, when the next surge occurs on the international agricultural markets, the international donor community should direct its resources towords countries in which poor households are particularly vulnerable to international food price spikes and where the government is unable to to provide sufficient assistance, rather than the LICs in general. This could be LICs which, for some reason, has to import most of its food goods. Secondly, the strong focus in the media and among agricultural economists on international food crises is misguided in the sense that LICs, in general, are primarily vulnerable to domestic rather than international shocks.