### 2015-12-22 regressions.Rnw
regression file for 1. round rev. v. of paper

% !Rnw root = Master.Rnw

\SweaveInput{data_prep.Rnw}

<<dom_infl_0708,echo=F>>=
library(sandwich)

m1 = lm(foodinfl78~ log(GDP)+I(log(GDP)^2)+ staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m1, type = "HC1")
rse1 = sqrt(diag(cov))

m2 = lm(foodinfl78~log(cons)+I(log(cons)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m2, type = "HC1")
rse2 = sqrt(diag(cov))

m3 = lm(foodinfl78~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m3, type = "HC1")
rse3 = sqrt(diag(cov))

m4 = lm(rfpigr78~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m4, type = "HC1")
rse4 = sqrt(diag(cov))

m5 = lm(rfpigr78~log(cons)+I(log(cons)^2) + staple_share*cdep,
       data = data_0708)
cov = vcovHC(m5, type = "HC1")
rse5 = sqrt(diag(cov))

m6 = lm(rfpigr78~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_0708)
cov = vcovHC(m6, type = "HC1")
rse6 = sqrt(diag(cov))
@

<<dom_infl_1011, echo=F>>=

m1 = lm(foodinfl1011~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m1, type = "HC1")
rse1 = sqrt(diag(cov))

m2 = lm(foodinfl1011~log(cons)+I(log(cons)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m2, type = "HC1")
rse2 = sqrt(diag(cov))

m3 = lm(foodinfl1011~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m3, type = "HC1")
rse3 = sqrt(diag(cov))

m4 = lm(rfpigr1011~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m4, type = "HC1")
rse4 = sqrt(diag(cov))

m5 = lm(rfpigr1011~log(cons)+I(log(cons)^2) + staple_share*cdep,
       data = data_1011)
cov = vcovHC(m5, type = "HC1")
rse5 = sqrt(diag(cov))

m6 = lm(rfpigr1011~log(food)+I(log(food)^2) + staple_share*cdep, 
       data = data_1011)
cov = vcovHC(m6, type = "HC1")
rse6 = sqrt(diag(cov))
@

<<ts_pt, echo=FALSE>>=
library(dynlm)

ptmat = matrix(, nrow = nrow(data_0708), ncol = 2)
data$id = as.numeric(data$iso3c)
for (i in 1:max(data$id)) {
  dat = data[data$id==i,]
  ptmat[i,1] = as.character(dat[1,"iso3c"])
  dat = ts(dat, start=c(2005, 1), end=c(2010, 12), frequency=12)
  m = dynlm(d(log(afpi))~ d(log(fpi)) + 
              L(d(log(fpi)),1:6) +
              L(d(log(afpi)),1:6), data = dat)
  ptmat[i,2] = sum(m$coefficients[2:8])/(1-sum(m$coefficients[9:14]))
}
ptdat = as.data.frame(ptmat)
names(ptdat) = c("iso3c", "lri")
ptdat = merge(ptdat, data_0708, by = "iso3c", all.x=TRUE)
ptdat$lri = as.numeric(ptdat$lri)

m1 = lm(lri~log(GDP) + I(log(GDP)^2)+staple_share*cdep, data = ptdat)
cov = vcovHC(m1, type = "HC1")
rse1 = sqrt(diag(cov))

m2 = lm(lri~log(cons) + I(log(cons)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m2, type = "HC1")
rse2 = sqrt(diag(cov))

m3 = lm(lri~log(food) + I(log(food)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m3, type = "HC1")
rse3 = sqrt(diag(cov))

ptmat = matrix(, nrow = nrow(data_0708), ncol = 2)
data$id = as.numeric(data$iso3c)
for (i in 1:max(data$id)) {
  dat = data[data$id==i,]
  ptmat[i,1] = as.character(dat[1,"iso3c"])
  dat = ts(dat, start=c(2005, 1), end=c(2010, 12), frequency=12)
  m = dynlm(d(log(rafpi))~ d(log(fpi)) + 
              L(d(log(fpi)),1:6) +
              L(d(log(rafpi)),1:6), data = dat)
  ptmat[i,2] = sum(m$coefficients[2:8])/(1-sum(m$coefficients[9:14]))
}
ptdat1 = as.data.frame(ptmat)
names(ptdat1) = c("iso3c", "lri1")
ptdat = merge(ptdat, ptdat1, by = "iso3c", all.x=TRUE)
ptdat$lri1 = as.numeric(ptdat$lri1)

m4 = lm(lri1~log(GDP)+I(log(GDP)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m4, type = "HC1")
rse4 = sqrt(diag(cov))

m5 = lm(lri1~log(cons) + I(log(cons)^2) + staple_share*cdep,
       data = ptdat)
cov = vcovHC(m5, type = "HC1")
rse5 = sqrt(diag(cov))

m6 = lm(lri1~log(food) + I(log(food)^2) + staple_share*cdep, 
       data = ptdat)
cov = vcovHC(m6, type = "HC1")
rse6 = sqrt(diag(cov))
@

<<panel_pt, echo=F>>=
library(plm)

m1 = plm(diff(log(afpi)) ~ lag(diff(log(afpi)),1:6)+
           lag(diff(log(fpi)),6)*log(GDP) +
           lag(diff(log(fpi)),6)*I(log(GDP)^2),
        data = data, model = "within")

cov = vcovHC(m1, method = "arellano", type = "HC1", cluster = "group")
rse1 = sqrt(diag(cov))

m2 = plm(diff(log(afpi)) ~ lag(diff(log(afpi)),1:6)+
           lag(diff(log(fpi)),6)*log(cons) +
           lag(diff(log(fpi)),6)*I(log(cons)^2),
        data = data, model = "within")

cov = vcovHC(m2, method = "arellano", type = "HC1", cluster = "group")
rse2 = sqrt(diag(cov))

m3 = plm(diff(log(afpi)) ~ lag(diff(log(afpi)),1:6)+
           lag(diff(log(fpi)),6)*log(food) +
           lag(diff(log(fpi)),6)*I(log(food)^2),
        data = data, model = "within")

cov = vcovHC(m3, method = "arellano", type = "HC1", cluster = "group")
rse3 = sqrt(diag(cov))

m4 = plm(diff(log(rafpi)) ~ lag(diff(log(rafpi)),1:6)+
           lag(diff(log(fpi)),6)*log(GDP) +
           lag(diff(log(fpi)),6)*I(log(GDP)^2),
        data = data, model = "within")

cov = vcovHC(m4, method = "arellano", type = "HC1", cluster = "group")
rse4 = sqrt(diag(cov))

m5 = plm(diff(log(rafpi)) ~ lag(diff(log(rafpi)),1:6)+
           lag(diff(log(fpi)),6)*log(cons) +
           lag(diff(log(fpi)),6)*I(log(cons)^2),
        data = data, model = "within")

cov = vcovHC(m5, method = "arellano", type = "HC1", cluster = "group")
rse5 = sqrt(diag(cov))

m6 = plm(diff(log(rafpi)) ~ lag(diff(log(rafpi)),1:6)+
           lag(diff(log(fpi)),6)*log(food) +
           lag(diff(log(fpi)),6)*I(log(food)^2),
        data = data, model = "within")

cov = vcovHC(m6, method = "arellano", type = "HC1", cluster = "group")
rse6 = sqrt(diag(cov))
@