### 2016-01-29 analysis.Rnw
Revised version

% !Rnw root = master1.Rnw

\section{Empirical analysis\label{sec:analysis}}
This section assesses the model's predictions regarding the relationship between income and price transmission. The model in section \ref{sec:model} predicts an inverse-U shaped relationship between income and price transmission but it does not pin down the actual turning point, which is an empirical problem. Our goal in this section is thus to a) establish whether there is a statistically significant quadratic relationship between income or, alternatively, expenditure and food price transmission and b) to determine the income/expenditure level corresponding to the turning point. Section \ref{sec:data} introduces the food price data and provides some context. Results from the statistical analysis are presented in section \ref{sec:analysis1} and \ref{sec:analysis2}.

\subsection{Data\label{sec:data}}

<<echo=FALSE>>=
load("infl_0708_1011.Rda")
load("ts_dat.Rda")
n.full = nrow(table(ts.dat$id))
n.sub = nrow(infl)
infl = infl[infl$iso3c != "BLR", ]
@

\subsubsection{Domestic and international food prices}
Domestic price data is sourced from the International Labor Organization (ILO).\footnote{See \citet{ilo16}} ILO maintains a comprehensive database of domestic food price indexes (FPIs) and consumer price indexes (CPIs) which is used to monitor the cost of living and real wages. There are not many studies which draw on this database, however, despite a broad country coverage and long monthly price series. A major issue with this data is that the historical series are not recalculated whenever a new base year is introduced. The same applies when other changes are made such as the inclusion or exclusion of certain items such as beverages or tobacco from the index. This means that many of the series have discrete level shifts associated with these changes. Another issue with the data is that some countries report several series representing different socioeconomic groups or geographical areas. In some cases one series replace another after a period of overlap. Finally, many of the series contain missing values. To ``tidy up'' the raw data I carried out the following steps First, I restricted the sample to the period 2005:1-2014:12. Secondly, in cases where multiple complete series were available from the same country, a single one was selected. Thirdly, series with missing values were discarded. Fourthly, and most critical for the analysis below, level shifts associated with base year changes etc. were identified and removed.\footnote{Concretely I added the change in the break point from the previous value to the break point and all points succeeding it. The value of the adjusted series at the break point is thus the same as the one preceding it.} Finally, all the series were normalized with respect to a common base period (Jan 2007). In total \Sexpr{n.full} countries have complete price series for the entire ten year period considered whereas, for example, there are \Sexpr{n.sub} countries with complete price series covering the 2005-11 subperiod. There is therefore a trade-off between the number of complete price series and the length of the series.

An important detail to know about this data is that many of the low income countries' price indexes derive from prices collected from one or two major markets in the capital. Furthermore, weights and selected items are derived from household expenditure surveys which often target certain groups such as the urban middle class. Differences in consumption patterns around the world of course translate into differences in weights and in the composition of the food baskets themselves. This is probably the largest drawback of using index data in a cross section analysis of price transmission. That being said, even domestic prices of individual food commodities such as wheat or rice would not refer to the same good due to quality and varietal differences, so it is not the case that focusing on a few selected food items consumed by all countries would necessarily solve the problem.

International food prices are represented by the FAO food price index (FAO FPI) and in some cases its ``real'' version, namely the nominal FAO FPI deflated by the Manufacturing Unit Value (MUV) index published by the World Bank.\footnote{see \citet{fao16} and \citet{worldbank16a}.} Table \ref{tab:ts_data} contain information about all time series data employed in the analysis below.

\subsubsection{Domestic food inflation during the food crises}

There has been two recent international price spikes; one in 2007-8 and the other in 2010-11. The former spike especially is often referred to as ``the food criss'' in the literature but there are also studies that consider the latter price spike \citep[e.g.][]{Headey2008, Headey2013, timmer10, Ivanic2012, Martin2011}. Between January 2007 and June 2008 FAO's food and grain price indexes rose by 91 and 127 points to index 226 and 268, respectively, corresponding to relative price increases of 68 and 90 percent. From the first peak in mid-2008 and to their their respective troughs in February and September 2009 the two indexes fell by 82 and 113 points before rising back up to index 240 and 261 in February and April 2011, respectively.

The boxplot in figure \ref{fig:infl} depicts the empirical distribution of the sample countries' headline (CPI) inflation, their food (FPI) inflation as well as the change in their FPI relative to the CPI, termed real FPI growth, over the two time periods 2007:1-2008:6 and 2010:1-2010:12.

\begin{center}
[Figure \ref{fig:infl} around here] \\
\par\end{center}

The median country experienced an annualized rate of food and headline inflation of \Sexpr{round(median(infl$fpigr.0708),1)} and \Sexpr{round(median(infl$cpigr.0708),1)} percent, respectively, in 2007-8. Nominal food prices are not only driven by exogenous world market shocks. They follow a stochastic trend together with the non-food prices, representing underlying general (core) inflation. By dividing the FPI with the CPI, we obtain an FPI which is purged of underlying domestic inflation. Real FPI growth, which is approximately the difference between the FPI and CPI inflation rate, was \Sexpr{round(median(infl$rfpigr.0708),1)} percent for the median country in that period.\footnote{A problem with the real FPI is that domestic non-food prices are also affected by international commodity shocks. Low or even negative growth in real domestic food prices in these two periods could therefore reflect a response to the concurrent surge in non-food commodities, oil in particular, rather than a puzzling negative response to an international food price shock.}

Price increases were generally lower in the second period. The median country experienced food and headline inflation rates \Sexpr{round(median(infl$fpigr.1011),1)} and \Sexpr{round(median(infl$cpigr.1011),1)} percent and a modest real FPI growth of \Sexpr{round(median(infl$rfpigr.1011),1)} percent. In both periods the variation in inflation rates is substantial. The standard deviation and interquartile range of 2007-8 food inflation is \Sexpr{round(sd(infl$fpigr.0708),1)} and \Sexpr{round(IQR(infl$fpigr.0708),1)} percentage points, respectively, versus \Sexpr{round(sd(infl$fpigr.1011),1)} and \Sexpr{round(IQR(infl$fpigr.1011),1)} percentage points in 2010-11. Although median real FPI growth is lower in the second period the variation is similar in both periods (standard deviations of \Sexpr{round(sd(infl$rfpigr.0708),1)} and \Sexpr{round(sd(infl$rfpigr.1011),1)} and interquartile ranges of \Sexpr{round(IQR(infl$rfpigr.0708),1)} and \Sexpr{round(IQR(infl$rfpigr.1011),1)}, respectively). The three remaining panels contain alternative representations the same data and there appears to be some systematic variation. As the (mostly) positively sloped lowess lines indicate, a country hit by a large nominal and real FPI shock in 2007-8 was more likely to be hit by a relatively large shock again in 2010-11 and vice versa. This suggests that there are indeed characteristics which influences a country's response to a world market shock, an assertion that will be explored further in section \ref{sec:analysis1} and \ref{sec:analysis2}.

\subsubsection{Country characteristics}
Table \ref{tab:cs_data} lists the variables that, based on the discussion in section \ref{sec:model}, are considered potentially important determinants of price transmission. I consider two proxies for a country's income level, namely country $i$'s (PPP adjusted) per capita GDP as well as its household consumption expenditure. The other covariates are a country $i$'s percentage dietary energy supply derived from cereals, roots and tubers, its cereal import- and export shares, a dummy indicating whether the country is a landlocked developing country and its trade openness defined as the sum of its imports and exports relative to its GDP. In addition there are two indexes included representing how easy it is to trade with the country in question and the quality of its infrastructure.

The analysis in section \ref{sec:model} suggests that international price shocks are transmitted the strongest to middle income countries so we would expect food price inflation in these two periods to follow an inverse-U shaped relationship with per capita income and consumption expenditure.

Some populations consume more basic food goods than others, for cultural reasons, which make their domestic food price level more responsive to price changes on grains and other internationally traded agricultural commodities. Furthermore, poor households also spend a larger share of their food budget on staples. Therefore, if we did account for this, the estimated relationship between income and domestic food price changes could be spurious.

Net food importers as well as exporters are, all else equal, more connected to the world market and domestic food prices in these countries should therefore react stronger to world market price shocks than prices in countries that are self sufficient in food. Countries that are largely dependent on grain imports to cover their domestic consumption needs can lower tariffs in response to an international price shock, but this will only have a substantial effect if tariffs are high to begin with. Many developing countries lowered their tariffs during the food crisis but the scope for major reductions was there only for a few countries \citet{imf08, Demeke2009}. All in all, we would expect to find a stronger transmission of international food price spikes to domestic price increases among countries with large import dependency ratios. Food importers could, of course, resort more often than other countries to domestic price stabilizing policies. In this case, there would be no or perhaps even a negative relationship between a country's import share and its price transmission rate. Such considerations also apply to the cereal consumption share discussed above.

Net food exporters, on the other hand, can use export restrictions as way of stabilizing domestic prices and some countries often do. Major grain exporters such as India, Argentina, Russia and Ukraine, among others, intervened in their export markets during the food crisis \citep{mitra09}. The question is whether Such interventions are, however, highly unpopular among the farmers and do not always work.

\subsection{Determinants of domestic food price changes during the 2007-8 and 2010-11 food crises\label{sec:analysis1}}
This section will discuss whether the cross country distribution of domestic food price changes during the recent international price spikes was as expected. That is, did middle income countries with open markets, good infrastructure and large shares of grains in their food consumption experience the greatest food price increases in these two periods? A discussion of whether there is a systematic variation in the degree to which domestic food prices react to international food price changes in general is postponed to section \ref{sec:analysis2}.

Table \ref{tab:fpi-regressions} and \ref{tab:rfpi-regressions} consider the cross sectional variation in domestic food price changes during the 2007-8 and 2010-11 food crises. Each of the columns refer to a regression of FPI inflation or real FPI growth on a subset of the explanatory variables discussed above. The models can all be written
\begin{equation}
\pi_{i}^{dom}=\delta_{0} + \delta_{1}x_{i} + u_{i}\label{eq:infl_model}
\end{equation}
where $\pi_{i}^{dom}$ refers to either country $i$'s FPI inflation or its real FPI growth in either the 18 month period 2007:1-2008:6 or the 12 month period 2010:1-2010:12. $x_{i}$ is the covariate vector and $u_{i}$ is the error term.

\begin{center}
[Table \ref{tab:fpi-regressions} around here]
\par\end{center}

In table \ref{tab:fpi-regressions}, the non-income related explanatory variables that are significant in half of the specifications or more are a country's cereal consumption share, its cereal import and export shares and landlocked status. As expected, countries with large cereal consumption shares experienced significantly higher food inflation than countries with lower cereal consumption shares during the 2010-11 international price spike, when we control for other factor affecting price transmission. The estimated effect is negative for the 2007-08 food crisis period but much smaller in magnitude and statistically insignificant. This could suggests that vulnerable countries intervened more heavily in their food markets during the first spike than in the second. LLDCs experienced significantly lower food inflation in the latter period (around 4 percentage points less on average) but not in the former. Again, the lack of significance could be a result of more intervention in the former period, which makes the relationship more noisy. Grain exporters experienced more food inflation on average in both periods but the effect is very small. Perhaps the most surprising finding is that grain importers experienced lower food inflation on average in both periods. As discussed above, this would suggest that large food importers intervene more in their food markets than other countries.

\begin{center}
[Table \ref{tab:rfpi-regressions} around here]
\par\end{center}

Table \ref{tab:rfpi-regressions} considers domestic real FPI growth rather than FPI inflation during during the 2007-08 and 2010-11 international price spikes. The main differences compared to table \ref{tab:fpi-regressions} is that countries with better infrastucture, i.e. a higher LPI index, are found to have higher real FPI growth rates but only statistically significantly so in the former period. The $R^2$ is quite high in all specifications, especially for the FPI inflation regressions in table \ref{tab:fpi-regressions}.\footnote{GDP and consumption expenditure contribute up to around half of the $R^2$ according to the LMG metric \citep{lindeman80}. These calculations were carried out in \textsf{R} using the \emph{relaimpo} package \citep{groemping06}.}

In order to find out whether the negative coefficient on the Cereal import variable and the insignificance of the Landlocked dummy in 2007-08 is due to policy interventions during the food crisis, I included a set of intervention dummies to the 2007-08 regressions in table \ref{tab:fpi-regressions} and \ref{tab:rfpi-regressions}. Results are shown in table \ref{tab:int-regressions}.\footnote{These dummies are based on table 1 in \citet{Demeke2009}. ``Export restrictions'' = 1 if the country Restricted or banned export, ``Price controls'' = 1 if the country administered price control or restricted private trade, ``Stock releases'' = 1 if the country released stocks (public or imported) at subsidized prices'', ``Tariff reductions'' = 1 if the country reduced tariffs or customs fees on imports and ``Tax reductions'' = 1 if the country suspended or reduced VAT or other taxes, all as of 1. December 2008.} As can be seen, this increases the coefficients on the LLDC dummy but not enough to achieve statistical significance at conventional levels. Furthermore, the coefficient on the Cereal import variable remains negative. Only the Tax reduction dummy is significant in more than one specification. The coefficient is positive though, suggesting that tax reductions did not lead to lower food prices increases during the food crisis, rather the opposite.

\subsection{Determinants of price transmission \label{sec:analysis2}}
Although the results in section \ref{sec:analysis1} provide insights into the cross sectional variation in domestic price changes during the two recent international food price spikes they have at least two shortcomings. First, none of the estimated relationships represent price transmission as such as there may be the other and perhaps more important causes of the observed domestic price changes. Secondly, the estimates are based on two specific time periods, which means that we ignore much the information contained in the individual price series and it also raises questions about the reliability of the results. To accommodate these shortcomings I approach the issue from two different angles. First, in line with the small existing literature on global food price transmission patterns, I estimate country specific price transmission coefficients based on a time series model. These estimated coefficients are then subsequently regressed on the same income and expenditure predictors as in section \ref{sec:analysis1}. Secondly, I estimate the effects of the predictor variables on price transmission directly in a panel model.

\subsubsection{Meta analysis of estimated price transmission rates\label{sec:meta-analysis}}
This section first presents a set of country specific price transmission estimates. Next a meta-regression analysis is carried out to see whether the estimates vary systematically with the proposed determinants of price transmission.

Following the literature, I regress monthly domestic food price changes on its lagged domestic food price changes as well as on contemporaneous and lagged international food price changes and exchange rates \citep[see][]{Ianchovichina2014, IMF2011, Kalkuhl2014}. That is, for each country $i$ an autoregressive distributive lag (ADL) models was fitted
\begin{equation}
\pi_{i,t}^{dom}=\delta_{i,0}+\sum_{s=1}^{p}\rho_{i,s}\pi_{i,t-s}^{dom}+\sum_{s=0}^{q}\lambda_{i,s}\pi_{t-s}^{int}+\sum_{s=0}^{q}\phi_{i,s}er_{t-s}+\epsilon_{i,t}\label{eq:time_series_model}
\end{equation}
where $\pi_{i,t}^{dom}$ represents country $i$'s monthly FPI inflation or real FPI growth at time $t$ (both approximated by the first differences to the log of FPI or real FPI), $\pi_{t}^{int}$ is the FAO food price index at time $t$ and $er_{i,t}$ is the country $i$'s domestic currency/US\$ exchange rate, both in log differences.

The ADL model is appropriate when the causal impact is unidirectional. This could be the case when the domestic price is from a small country relative to the world market but, more often, this assumption is invoked when the domestic price variable is a Food Price Index (FPI), i.e. the food component of the Consumer Price Index (CPI).

Prior to estimation I tested each of the series for unit roots and cointegration with international prices.\footnote{I tested for stationarity with the KPSS test \citep{kwiatkowski92} and for unit roots with the augmented Dickey-Fuller test \citep{dickey79}. I used the Phillips and Ouliaris ``Pz'' test \citep{phillips90} to test for cointegration. These tests are implemented in \textsf{R} via the packages \emph{tseries} \citep{trapletti15} and \emph{urca} \citep{pfaff08}, respectively.} Results indicated that the domestic FPI and real FPI series, are nonstationary in levels but stationary in first differences. Domestic price levels were also found not to be cointegrated with world prices. Consequently, each model was estimated in growth rates rather than levels in order to avoid the problems associated with spurious regressions.

Next, for each country, I calculated the long run multiplier (LRM) associated with a change in the the FAO food price index given by
\begin{equation}
\theta_{i}=\frac{\sum_{s=0}^{q}\lambda_{i,s}}{1-\sum_{s=1}^{p}\rho_{i,s}}.
\end{equation}

<<echo=FALSE>>=
library(dplyr)
load("lrm_fpi_l1_12.Rda")
fpi = select(lrm.fpi.l.1_12, -iso3c)
load("lrm_rfpi_l1_12.Rda")
rfpi = select(lrm.rfpi.l.1_12, -iso3c)
fivenum.fpi = data.frame(sapply(fpi, summary))
fivenum.rfpi = data.frame(sapply(rfpi, summary))
n = nrow(fpi)
@

The LRM is a common measure of price transmission. \cite{IMF2011}, for example, contains an analysis of food price price transmission based on a dataset of monthly FPI inflation from 31 advanced economies and 47 emerging and developing economies over the period 2000 - 2011. Their median estimated LRM pass-through of a 1 percent increase in international food inflation is 0.18 percent in developed countries and 0.34 percent in emerging and developing countries. Similarly, \citet{Ianchovichina2014} report an average estimated LRM for the eighteen Middle East and North Africa (MENA) countries of 0.25 and the majority of the countries are reported to have  food price pass-through LRMs in the order of 0.2-0.4. Both studies use 12 lags in their country specific ADL models, i.e. they assume $p=q=12$.\footnote{\cite{IMF2011} do not include exchange rates in their regressions.} My own estimates based on a sample of \Sexpr{n} countries covering the period 2005:1 to 2014:12 are very similar. The average and median FPI transmission LRM from a model with 12 lags of all variables are \Sexpr{round(fivenum.fpi["Mean", "V12"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V12"], 2)}, respectively. That being said, it is not obvious that so many lags should be included in the models. When lag length selection is based on the Akaike Information Criterion (AIC), for example, the optimal number lags in most of the models turn out to be $p=q=1$.\footnote{For each country, I calculated the AIC for each of the $12^2$ combinations of $(p,q)\, \in (1,..,12)$. Then I selected the $(p,q)$ combination with the lowest AIC.} But when a common lag length of 1 is used for all countries, average and median FPI transmission LRMs are much lower, namely \Sexpr{round(fivenum.fpi["Mean", "V1"], 2)} and \Sexpr{round(fivenum.fpi["Median", "V1"], 2)}, respectively.\footnote{I also tested for residual autocorrelation with the Durbin-Watson test. There was no evidence of residual autocorrelation in either case.}

Finally, FPI and Real FPI transmission estimates, $\hat{\theta}$, are regressed, on the same explanatory variables as in table \ref{tab:fpi-regressions} and \ref{tab:rfpi-regressions}. That is, the following model is fitted
\begin{equation}
\hat{\theta}_{i} =\delta_{0} + \delta_{1}x_{i} + u_{i}\label{eq:ptreg},
\end{equation}
where the covariates $x$ are the same as in equation (\ref{eq:infl_model}). The dependent variable in Equation (\ref{eq:ptreg}) is an estimate rather than an observed value and the the equation therefore represents an estimated dependent variable (EDV) regression \citep[see][]{hanushek74, lewis05}. A common approach in this situation is to use Weighted Least Squares (WLS) estimation with weights $w_{i}$ that are inversely proportional to the standard error of $\theta_{i}$ \citep{saxonhouse76}. This way more precise estimates are given a higher weight in the regression. The main reason though for using WLS rather than OLS is that the latter may suffer from heteroskedasticity which leads to inefficient coefficient estimates inconsistent standard errors. The inconsistent OLS standard errors can, however, easily be corrected with a robust standard error estimator. \citet{lewis05} uses formal reasoning backed up by a Monte Carlo analysis to argue that OLS with robust standard errors is likely to produce better estimates of parameter uncertainty than WLS based on inverse standard error weights, but the latter is likely to produce more efficient parameter estimates. Furthermore, they conclude that OLS with robust standard errors is the best approach unless information about the sampling error of the dependent variable is not only available but also highly reliable.

Table \ref{tab:fpi-pt-regressions} presents the estimates from equation \ref{eq:ptreg} based on OLS with robust standard errors as well as WLS. I tested for heteroskedasticity with the Breusch-Pagan test \citep{breusch79} but there was no evidence against the null of homoskedasticity. Consequently, ordinary and robust standard errors were very similar. The results also turned out to be robust to the choice of lag lengths in connection with the LRM and therefore, only the results based on the single lag LRM are reported.

\begin{center}
[Table \ref{tab:fpi-pt-regressions} around here]
\par\end{center}

As can be seen, results are fairly robust to the choice of estimator although the significance levels are somewhat higher in the OLS case. As in the domestic food inflation case (table \ref{tab:fpi-regressions} -\ref{tab:int-regressions}) we find that cereal exporters respond stronger to international food price changes whereas cereal importers and cereal consumers respond less to internstional price shocks. These effects are, however, very small and/or statistically insignificant. LLDCs, curiously, seem to respond stronger on average to international price shocks than other countries. GDP and consumption expenditure are also apparently important determinants of food price transmission.

Finally, table \ref{tab:rfpi-pt-regressions} repeats the exercise for the LRM based on the real FPI data. Compared to table \ref{tab:fpi-pt-regressions}, not much changes.

\begin{center}
[Table \ref{tab:rfpi-pt-regressions} around here]
\par\end{center}

\subsubsection{A panel model of price transmission determinants}
The findings discussed above suggest that middle income countries are the most vulnerable to international food price spikes even after controlling for other factors which plausibly affect price transmission. However, the estimates in table \ref{tab:fpi-pt-regressions} and \ref{tab:rfpi-pt-regressions} are potentially subject to unobserved heterogeneity bias. As a last robustness check I therefore exploit the panel structure of the dataset to control for time invariant unobserved heterogeneity which potentially confounds the relationship between price transmission and its determinants. Specifically, the following model is fitted
\begin{equation}
\pi_{it}^{dom}=\delta_{0}+
\sum_{s=1}^{6}\rho_{s}\pi_{it-s}^{dom}+
\sum_{s=0}^{1}(\lambda_{s}\pi_{t-s}^{int}+\phi_{s}er_{it-s})+
\delta_{1}\pi_{t}^{int}\cdot x_{i} +
u_{it}\label{eq:panel_model},
\end{equation}
where all variables are the same as before. I tried to include more lags of $\pi_{t}^{int}$ and $er_{it}$ but their coefficients were small and generally insignificant. The main effect of $\pi^{int}$ on $\pi_{i}^{dom}$ is on impact, i.e. at time $t=0$. This is the reason why the interactions terms $\pi_{t}^{int}\cdot x_{i} $ also refer to the impact effect.

It is well known that witin estimation of dynamic panel models such as equation (\ref{eq:panel_model}) suffers from Nickell bias. The problem is that the demeaned lagged dependent variable is correlated with the demeaned error term which makes the estimator inconsistent \citep{nickell81}. This inconsistency, however, is mainly a problem in short panel datasets where the number of cross section obeservations $N$  is much larger than the number of time periods $T$. Specifically, \citet{nickell81} demonstrates that the bias is of the order $1/T$ as $N\rightarrow \infty$. In the present case where $T>100$ the bias should be negligible. Nontheless, as a last robustness check, I also consider the  Anderson-Hsiao (AH) estimator \citep{anderson82} based on the first-difference transformation of the model in log levels with lags 7-12 of the dependent variable acting as instruments.

Estimation results are shown in table \ref{tab:fe-panel-regressions}-\ref{tab:fd-panel-regressions}.\footnote{Note that coefficients on the lagged dependent variable are omitted.} In table \ref{tab:fe-panel-regressions}, the only explanatory varibles that remain significant are those related to income, namely the GDP and Consumption expenditure terms. The effect sizes in table \ref{tab:fd-panel-regressions} are similar to those in table \ref{tab:fd-panel-regressions} but the standard errors are generally larger so the significance mostly disappears. Only the Cereal export share becomes significant in the FPI regressions.

<<echo=FALSE>>=
load("ctr_dat_0708.Rda")
dat = ctr.dat.0708
r1 = readRDS("fpi_reg_m.rds")
r2 = readRDS("rfpi_reg_m.rds")
r3 = readRDS("int_reg_m.rds")
r4 = readRDS("fpi_pt_reg_m.rds")
r5 = readRDS("rfpi_pt_reg_m.rds")
r6 = readRDS("panel_reg_m.rds")
r7 = readRDS("panel_reg_fd_m.rds")

tp = matrix(NA, nrow=4, ncol=7)

tp[1,1] = -coef(r1[[1]])["log(gdp)"]/(2*coef(r1[[1]])["I(log(gdp)^2)"])
tp[2,1] = -coef(r1[[2]])["log(cons)"]/(2*coef(r1[[2]])["I(log(cons)^2)"])
tp[3,1] = -coef(r1[[3]])["log(gdp)"]/(2*coef(r1[[3]])["I(log(gdp)^2)"])
tp[4,1] = -coef(r1[[4]])["log(cons)"]/(2*coef(r1[[4]])["I(log(cons)^2)"])

tp[1,2] = -coef(r2[[1]])["log(gdp)"]/(2*coef(r2[[1]])["I(log(gdp)^2)"])
tp[2,2] = -coef(r2[[2]])["log(cons)"]/(2*coef(r2[[2]])["I(log(cons)^2)"])
tp[3,2] = -coef(r2[[3]])["log(gdp)"]/(2*coef(r2[[3]])["I(log(gdp)^2)"])
tp[4,2] = -coef(r2[[4]])["log(cons)"]/(2*coef(r2[[4]])["I(log(cons)^2)"])

tp[1,3] = -coef(r3[[1]])["log(gdp)"]/(2*coef(r3[[1]])["I(log(gdp)^2)"])
tp[2,3] = -coef(r3[[2]])["log(cons)"]/(2*coef(r3[[2]])["I(log(cons)^2)"])
tp[3,3] = -coef(r3[[3]])["log(gdp)"]/(2*coef(r3[[3]])["I(log(gdp)^2)"])
tp[4,3] = -coef(r3[[4]])["log(cons)"]/(2*coef(r3[[4]])["I(log(cons)^2)"])

tp[1,4] = -coef(r4[[1]])["log(gdp)"]/(2*coef(r4[[1]])["I(log(gdp)^2)"])
tp[2,4] = -coef(r4[[2]])["log(cons)"]/(2*coef(r4[[2]])["I(log(cons)^2)"])
tp[3,4] = -coef(r4[[3]])["log(gdp)"]/(2*coef(r4[[3]])["I(log(gdp)^2)"])
tp[4,4] = -coef(r4[[4]])["log(cons)"]/(2*coef(r4[[4]])["I(log(cons)^2)"])

tp[1,5] = -coef(r5[[1]])["log(gdp)"]/(2*coef(r5[[1]])["I(log(gdp)^2)"])
tp[2,5] = -coef(r5[[2]])["log(cons)"]/(2*coef(r5[[2]])["I(log(cons)^2)"])
tp[3,5] = -coef(r5[[3]])["log(gdp)"]/(2*coef(r5[[3]])["I(log(gdp)^2)"])
tp[4,5] = -coef(r5[[4]])["log(cons)"]/(2*coef(r5[[4]])["I(log(cons)^2)"])

tp[1,6] = -coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
  (2*coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
tp[2,6] = -coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
  (2*coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])
tp[3,6] = -coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
  (2*coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
tp[4,6] = -coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
  (2*coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])

tp[1,7] = -coef(r7[[1]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
  (2*coef(r7[[1]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
tp[2,7] = -coef(r7[[2]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
  (2*coef(r7[[2]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])
tp[3,7] = -coef(r7[[3]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
  (2*coef(r7[[3]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
tp[4,7] = -coef(r7[[4]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
  (2*coef(r7[[4]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])

tp = exp(tp)

nc = matrix(NA, nrow=4, ncol=7)
nc.pc = matrix(NA, nrow=4, ncol=7)
n.gdp = nrow(dat[which(dat$gdp>0), ])
n.cons = nrow(dat[which(dat$cons>0), ])

nc[1,1] = nrow(dat[which(dat$gdp< tp[1,1] & dat$gdp >0), ])
nc[2,1] = nrow(dat[which(dat$cons< tp[2,1] & dat$cons >0), ])
nc[3,1] = nrow(dat[which(dat$gdp< tp[3,1] & dat$gdp >0), ])
nc[4,1] = nrow(dat[which(dat$cons< tp[4,1] & dat$cons >0), ])

nc.pc[1,1] = round(nc[1,1]/n.gdp*100,0)
nc.pc[2,1] = round(nc[2,1]/n.cons*100,0)
nc.pc[3,1] = round(nc[3,1]/n.gdp*100,0)
nc.pc[4,1] = round(nc[4,1]/n.cons*100,0)

nc[1,2] = nrow(dat[which(dat$gdp< tp[1,2] & dat$gdp >0), ])
nc[2,2] = nrow(dat[which(dat$cons< tp[2,2] & dat$cons >0), ])
nc[3,2] = nrow(dat[which(dat$gdp< tp[3,2] & dat$gdp >0), ])
nc[4,2] = nrow(dat[which(dat$cons< tp[4,2] & dat$cons >0), ])

nc.pc[1,2] = round(nc[1,2]/n.gdp*100,0)
nc.pc[2,2] = round(nc[2,2]/n.cons*100,0)
nc.pc[3,2] = round(nc[3,2]/n.gdp*100,0)
nc.pc[4,2] = round(nc[4,2]/n.cons*100,0)

nc[1,3] = nrow(dat[which(dat$gdp< tp[1,3] & dat$gdp >0), ])
nc[2,3] = nrow(dat[which(dat$cons< tp[2,3] & dat$cons >0), ])
nc[3,3] = nrow(dat[which(dat$gdp< tp[3,3] & dat$gdp >0), ])
nc[4,3] = nrow(dat[which(dat$cons< tp[4,3] & dat$cons >0), ])

nc.pc[1,3] = round(nc[1,3]/n.gdp*100,0)
nc.pc[2,3] = round(nc[2,3]/n.cons*100,0)
nc.pc[3,3] = round(nc[3,3]/n.gdp*100,0)
nc.pc[4,3] = round(nc[4,3]/n.cons*100,0)

nc[1,4] = nrow(dat[which(dat$gdp< tp[1,4] & dat$gdp >0), ])
nc[2,4] = nrow(dat[which(dat$cons< tp[2,4] & dat$cons >0), ])
nc[3,4] = nrow(dat[which(dat$gdp< tp[3,4] & dat$gdp >0), ])
nc[4,4] = nrow(dat[which(dat$cons< tp[4,4] & dat$cons >0), ])

nc.pc[1,4] = round(nc[1,4]/n.gdp*100,0)
nc.pc[2,4] = round(nc[2,4]/n.cons*100,0)
nc.pc[3,4] = round(nc[3,4]/n.gdp*100,0)
nc.pc[4,4] = round(nc[4,4]/n.cons*100,0)

nc[1,5] = nrow(dat[which(dat$gdp< tp[1,5] & dat$gdp >0), ])
nc[2,5] = nrow(dat[which(dat$cons< tp[2,5] & dat$cons >0), ])
nc[3,5] = nrow(dat[which(dat$gdp< tp[3,5] & dat$gdp >0), ])
nc[4,5] = nrow(dat[which(dat$cons< tp[4,5] & dat$cons >0), ])

nc.pc[1,5] = round(nc[1,5]/n.gdp*100,0)
nc.pc[2,5] = round(nc[2,5]/n.cons*100,0)
nc.pc[3,5] = round(nc[3,5]/n.gdp*100,0)
nc.pc[4,5] = round(nc[4,5]/n.cons*100,0)

nc[1,6] = nrow(dat[which(dat$gdp< tp[1,6] & dat$gdp >0), ])
nc[2,6] = nrow(dat[which(dat$cons< tp[2,6] & dat$cons >0), ])
nc[3,6] = nrow(dat[which(dat$gdp< tp[3,6] & dat$gdp >0), ])
nc[4,6] = nrow(dat[which(dat$cons< tp[4,6] & dat$cons >0), ])

nc.pc[1,6] = round(nc[1,6]/n.gdp*100,0)
nc.pc[2,6] = round(nc[2,6]/n.cons*100,0)
nc.pc[3,6] = round(nc[3,6]/n.gdp*100,0)
nc.pc[4,6] = round(nc[4,6]/n.cons*100,0)

nc[1,7] = nrow(dat[which(dat$gdp< tp[1,7] & dat$gdp >0), ])
nc[2,7] = nrow(dat[which(dat$cons< tp[2,7] & dat$cons >0), ])
nc[3,7] = nrow(dat[which(dat$gdp< tp[3,7] & dat$gdp >0), ])
nc[4,7] = nrow(dat[which(dat$cons< tp[4,7] & dat$cons >0), ])

nc.pc[1,7] = round(nc[1,7]/n.gdp*100,0)
nc.pc[2,7] = round(nc[2,7]/n.cons*100,0)
nc.pc[3,7] = round(nc[3,7]/n.gdp*100,0)
nc.pc[4,7] = round(nc[4,7]/n.cons*100,0)

nc.pc.colmean = colMeans(nc.pc)
nc.pc.mean = mean(nc.pc.colmean)
nc.pc.panel.mean = mean(nc.pc.colmean[6:7])
nc.pc.nonpanel.mean = mean(nc.pc.colmean[-(6:7)])
@

\subsection{Estimated income turning points}
Table \ref{tab:tp} summarizes the estimated income turning points derived from the coefficients in table \ref{tab:fpi-regressions}-\ref{tab:fd-panel-regressions}. The figures in the last column refer to the number of countries with a per capita GDP or household consumption expenditure less than the estimated turning point shown in the third-last column. The average number of countries that are to the left of the turning point relative to the number of countries in the sample with data on these variables, across all specifications, is \Sexpr{round(nc.pc.mean,0)} percent. There are, however, important differences between the different estimators. Panel estimation, in particular, leads to a significantly higher proportion of countries to the left of the turning point than cross section estimation (\Sexpr{round(nc.pc.panel.mean,0)} vs. \Sexpr{round(nc.pc.nonpanel.mean,0)} percent) on average.