### 2016-01-29 tables.Rnw
Revised version. Note that the tp table created a failed message.

% !Rnw root = master1.Rnw

\section{Tables}

\begin{table}[h]
\caption{Time series data on prices and exchange rates}
\label{tab:ts_data}
\begin{center}
{\small{}}%
\begin{tabular}{|c|>{\raggedright}m{5cm}|>{\raggedright}m{3cm}|>{\raggedright}p{3cm}|}
\hline 
\textbf{\small{}Variable} & \textbf{\small{}Description} & \textbf{\small{}Source} & \textbf{\small{}Expected effect}\tabularnewline
\hline 
{\small{}FPI} & {\small{}Domestic food price index.} & {\small{}ILO, LABORSTA} & -\tabularnewline
\hline
{\small{}CPI} & {\small{}Domestic consumer price index.} & {\small{}ILO, LABORSTA} & -\tabularnewline
\hline 
{\small{}Real FPI} & {\small{}Domestic real food price index. FPI deflated by the CPI} & {\small{}ILO, LABORSTA} & -\tabularnewline
\hline 
{\small{}FAO FPI} & {\small{}FAO food price index.} & {\small{}FAO} & Positive\tabularnewline
\hline 
{\small{}Real FAO FPI} & {\small{}Real FAO food price index. Nominal index deflated by the
World Bank Manufacturing Unit Value index} & {\small{}FAO} & Positive\tabularnewline
\hline 
{\small{}Exchange rate} & {\small{}Domestic currency/US\$ exchange rate} & {\small{}IMF} & Positive\tabularnewline
\hline 
\end{tabular}{\small \par}
\end{center}
\end {table}

\begin{table}[h]
\caption{Cross sectional data on country characteristics}
\label{tab:cs_data}
\begin{center}

{\small{}}%
\begin{tabular}{|c|>{\raggedright}m{5cm}|>{\raggedright}m{3cm}|>{\raggedright}p{3cm}|}
\hline 
\textbf{\small{}Variable} & \textbf{\small{}Description} & \textbf{\small{}Source} & \textbf{\small{}Expected effect}\tabularnewline
\hline 
{\small{}GDP} & {\small{}Gross domestic product. Log of per capita GDP (current int\$, PPP)} & {\small{}International Comparison Program 2005. World Bank database} & Inverse-U\tabularnewline
\hline 
{\small{}Consumption} & {\small{}Household consumption. Log of individual consumption expenditure by households (current int\$, PPP)} & {\small{}International Comparison Program 2005. World Bank database} & Inverse-U\tabularnewline
\hline 
{\small{}Cereal consumption share} & {\small{}Share of dietary energy supply derived from
cereals, roots and tubers (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} & Positive\tabularnewline
\hline 
{\small{}Cereal import share} & {\small{}Cereal imports as a share of domestic supply (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} & Positive\tabularnewline
\hline 
{\small{}Cereal export share} & {\small{}Cereal exports as a share of domestic production (\%). 3-year average, 2004-06} & {\small{}FAO, FAOSTAT} & Positive\tabularnewline
\hline 
{\small{}LLDC} & {\small{}Dummy. 1 if country is a landlocked developing country} & {\small{}UNCTAD} & Negative\tabularnewline
\hline 
{\small{}Trade openness} & {\small{}Sum of exports and imports of goods and services as a share of GDP (\%). 2006-10 average} & {\small{}World Bank Development Indicators} & Positive\tabularnewline
\hline 
{\small{}DTF} & {\small{}Distance to Frontier index. An economy\textquoteright s distance to frontier measured on a scale from 0 to 100, where 0 represents the lowest performance and 100 the frontier.} & {\small{}Doing Business survey, Trading Across Borders . World Bank} & Positive\tabularnewline
\hline 
{\small{}LPI} & {\small{} Logistics performance index. The LPI measures ease of trade and transport of goods. The index ranges from 1 to 5, with a higher score representing better performance.} & {\small{}Logistics Performance Index 2007. World Bank database} & Positive\tabularnewline
\hline 
\end{tabular}{\small \par}
\end{center}
\end {table}

<<results=tex,echo=F>>=
library(stargazer)
reg.list = readRDS("fpi_reg_m.rds")
se.list = readRDS("fpi_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic FPI inflation 2007-8 and 2010-11', 
          label = "tab:fpi-regressions", 
          dep.var.labels=c("FPI inflation 07-08", "FPI inflation 10-11"), 
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("log of GDP",
                             "log of GDP squared", 
                             "log of consumption", 
                             "log of consumption squared", 
                             "Cereal consumption share", 
                             "Cereal import share", 
                             "Cereal export share",
                             "LLDC",
                             "State trading enterprice",
                             "Trade openness", 
                             "DTF", 
                             "LPI"))
@

<<results=tex,echo=F>>=
reg.list = readRDS("rfpi_reg_m.rds")
se.list = readRDS("rfpi_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic real FPI growth 2007-8 and 2010-11', 
          label = "tab:rfpi-regressions", 
          dep.var.labels=c("Real FPI growth 07-08", "Real FPI growth 10-11"), 
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("log of GDP",
                             "log of GDP squared", 
                             "log of consumption", 
                             "log of consumption squared", 
                             "Cereal consumption share", 
                             "Cereal import share", 
                             "Cereal export share",
                             "LLDC",
                             "State trading enterprice", 
                             "Trade openness", 
                             "DTF", 
                             "LPI"))
@

<<results=tex,echo=F>>=
reg.list = readRDS("int_reg_m.rds")
se.list = readRDS("int_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic FPI inflation 2007-8', 
          label = "tab:int-regressions", 
          dep.var.labels=c("FPI inflation 07-08", "Real FPI growth 07-08"), 
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("log of GDP",
                             "log of GDP squared", 
                             "log of consumption", 
                             "log of consumption squared", 
                             "Cereal consumption share", 
                             "Cereal import share", 
                             "Cereal export share",
                             "LLDC", 
                             "State trading enterprice",
                             "Trade openness", 
                             "DTF", 
                             "LPI",
                             "Export restrictions",
                             "Price controls",
                             "Stock releases",
                             "Tariff reductions",
                             "Tax reductions"))
@

<<results=tex,echo=F>>=
library(stargazer)
reg.list = readRDS("fpi_pt_reg_m.rds")
se.list = readRDS("fpi_pt_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run FPI transmission rates', 
          label = "tab:fpi-pt-regressions", 
          dep.var.labels=c("FPI transmission LRM"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("log of GDP",
                             "log of GDP squared", 
                             "log of consumption", 
                             "log of consumption squared",
                             "Cereal consumption share", 
                             "Cereal import share", 
                             "Cereal export share",
                             "LLDC", 
                             "Trade openness", 
                             "DTF", 
                             "LPI"))
@

<<results=tex,echo=F>>=
reg.list = readRDS("rfpi_pt_reg_m.rds")
se.list = readRDS("rfpi_pt_reg_se.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run real FPI transmission rates', 
          label = "tab:rfpi-pt-regressions", 
          dep.var.labels=c("Real FPI transmission LRM"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize", 
          covariate.labels=c("log of GDP",
                             "log of GDP squared", 
                             "log of consumption", 
                             "log of consumption squared",
                             "Cereal consumption share", 
                             "Cereal import share", 
                             "Cereal export share",
                             "LLDC", 
                             "Trade openness", 
                             "DTF", 
                             "LPI"))
@

<<results=tex, echo=FALSE>>=
library(plm)
reg.list = readRDS("panel_reg_m.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Dynamic panel regressions (within (FE) estimator)', 
          label = "tab:fe-panel-regressions", 
          dep.var.labels=c("FPI inflation", "Real FPI growth"), 
          no.space=TRUE,
          omit = c("adj.fpi", "rfpi"),
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize",
          covariate.labels=c("$\\pi_{t}^{i}$",
                             "$\\pi_{t-1}^{i}$",
                             "$\\Delta \\log(er_{t})$",
                             "$\\Delta \\log(er_{t-1})$",
                             "$\\pi_{t}^{i}\\times \\text{log of GDP}$",
                             "$\\pi_{t}^{i}\\times \\text{log of GDP squared}$", 
                             "$\\pi_{t}^{i}\\times \\text{log of consumption}$", 
                             "$\\pi_{t}^{i}\\times \\text{log of consumption squared}$", 
                             "$\\pi_{t}^{i}\\times \\text{Cereal consumption share}$", 
                             "$\\pi_{t}^{i}\\times \\text{Cereal import share}$", 
                             "$\\pi_{t}^{i}\\times \\text{Cereal export share}$",
                             "$\\pi_{t}^{i}\\times \\text{LLDC}$", 
                             "$\\pi_{t}^{i}\\times \\text{Trade openness}$", 
                             "$\\pi_{t}^{i}\\times \\text{DTF}$", 
                             "$\\pi_{t}^{i}\\times \\text{LPI}$"))
@

<<results=tex, echo=FALSE>>=
library(plm)
reg.list = readRDS("panel_reg_fd_m.rds")

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Dynamic panel regressions (Anderson-Hsiao (AH) estimator)', 
          label = "tab:fd-panel-regressions", 
          dep.var.labels=c("FPI inflation", "Real FPI growth"), 
          no.space=TRUE,
          omit = c("adj.fpi", "rfpi"),
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "footnotesize",
          covariate.labels=c("$\\pi_{t}^{i}$",
                             "$\\pi_{t-1}^{i}$",
                             "$\\Delta \\log(er_{t})$",
                             "$\\Delta \\log(er_{t-1})$",
                             "$\\pi_{t}^{i}\\times \\text{log of GDP}$",
                             "$\\pi_{t}^{i}\\times \\text{log of GDP squared}$", 
                             "$\\pi_{t}^{i}\\times \\text{log of consumption}$", 
                             "$\\pi_{t}^{i}\\times \\text{log of consumption squared}$", 
                             "$\\pi_{t}^{i}\\times \\text{Cereal consumption share}$", 
                             "$\\pi_{t}^{i}\\times \\text{Cereal import share}$", 
                             "$\\pi_{t}^{i}\\times \\text{Cereal export share}$",
                             "$\\pi_{t}^{i}\\times \\text{LLDC}$", 
                             "$\\pi_{t}^{i}\\times \\text{Trade openness}$", 
                             "$\\pi_{t}^{i}\\times \\text{DTF}$", 
                             "$\\pi_{t}^{i}\\times \\text{LPI}$"))
@

<<echo=FALSE>>=

## Remember to change variable names!!!!!

load("ctr_dat_0708.Rda")
dat = ctr.dat.0708
r1 = readRDS("fpi_reg_m.rds")
r2 = readRDS("rfpi_reg_m.rds")
r3 = readRDS("int_reg_m.rds")
r4 = readRDS("fpi_pt_reg_m.rds")
r5 = readRDS("rfpi_pt_reg_m.rds")
r6 = readRDS("panel_reg_m.rds")
r7 = readRDS("panel_reg_fd_m.rds")

tp = matrix(NA, nrow=4, ncol=7)

tp[1,1] = -coef(r1[[1]])["log(gdp)"]/(2*coef(r1[[1]])["I(log(gdp)^2)"])
tp[2,1] = -coef(r1[[2]])["log(cons)"]/(2*coef(r1[[2]])["I(log(cons)^2)"])
tp[3,1] = -coef(r1[[3]])["log(gdp)"]/(2*coef(r1[[3]])["I(log(gdp)^2)"])
tp[4,1] = -coef(r1[[4]])["log(cons)"]/(2*coef(r1[[4]])["I(log(cons)^2)"])

tp[1,2] = -coef(r2[[1]])["log(gdp)"]/(2*coef(r2[[1]])["I(log(gdp)^2)"])
tp[2,2] = -coef(r2[[2]])["log(cons)"]/(2*coef(r2[[2]])["I(log(cons)^2)"])
tp[3,2] = -coef(r2[[3]])["log(gdp)"]/(2*coef(r2[[3]])["I(log(gdp)^2)"])
tp[4,2] = -coef(r2[[4]])["log(cons)"]/(2*coef(r2[[4]])["I(log(cons)^2)"])

tp[1,3] = -coef(r3[[1]])["log(gdp)"]/(2*coef(r3[[1]])["I(log(gdp)^2)"])
tp[2,3] = -coef(r3[[2]])["log(cons)"]/(2*coef(r3[[2]])["I(log(cons)^2)"])
tp[3,3] = -coef(r3[[3]])["log(gdp)"]/(2*coef(r3[[3]])["I(log(gdp)^2)"])
tp[4,3] = -coef(r3[[4]])["log(cons)"]/(2*coef(r3[[4]])["I(log(cons)^2)"])

tp[1,4] = -coef(r4[[1]])["log(gdp)"]/(2*coef(r4[[1]])["I(log(gdp)^2)"])
tp[2,4] = -coef(r4[[2]])["log(cons)"]/(2*coef(r4[[2]])["I(log(cons)^2)"])
tp[3,4] = -coef(r4[[3]])["log(gdp)"]/(2*coef(r4[[3]])["I(log(gdp)^2)"])
tp[4,4] = -coef(r4[[4]])["log(cons)"]/(2*coef(r4[[4]])["I(log(cons)^2)"])

tp[1,5] = -coef(r5[[1]])["log(gdp)"]/(2*coef(r5[[1]])["I(log(gdp)^2)"])
tp[2,5] = -coef(r5[[2]])["log(cons)"]/(2*coef(r5[[2]])["I(log(cons)^2)"])
tp[3,5] = -coef(r5[[3]])["log(gdp)"]/(2*coef(r5[[3]])["I(log(gdp)^2)"])
tp[4,5] = -coef(r5[[4]])["log(cons)"]/(2*coef(r5[[4]])["I(log(cons)^2)"])

tp[1,6] = -coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
  (2*coef(r6[[1]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
tp[2,6] = -coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
  (2*coef(r6[[2]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])
tp[3,6] = -coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * log(gdp))"]/
  (2*coef(r6[[3]])["I(lag(diff(log(fao.fpi)), 0) * I(log(gdp)^2))"])
tp[4,6] = -coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * log(cons))"]/
  (2*coef(r6[[4]])["I(lag(diff(log(fao.fpi)), 0) * I(log(cons)^2))"])

tp[1,7] = -coef(r7[[1]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
  (2*coef(r7[[1]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
tp[2,7] = -coef(r7[[2]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
  (2*coef(r7[[2]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])
tp[3,7] = -coef(r7[[3]])["I(lag(log(fao.fpi), 0) * log(gdp))"]/
  (2*coef(r7[[3]])["I(lag(log(fao.fpi), 0) * I(log(gdp)^2))"])
tp[4,7] = -coef(r7[[4]])["I(lag(log(fao.fpi), 0) * log(cons))"]/
  (2*coef(r7[[4]])["I(lag(log(fao.fpi), 0) * I(log(cons)^2))"])

tp = exp(tp)

nc = matrix(NA, nrow=4, ncol=7)
nc.pc = matrix(NA, nrow=4, ncol=7)
n.gdp = nrow(dat[which(dat$gdp>0), ])
n.cons = nrow(dat[which(dat$cons>0), ])

nc[1,1] = nrow(dat[which(dat$gdp< tp[1,1] & dat$gdp >0), ])
nc[2,1] = nrow(dat[which(dat$cons< tp[2,1] & dat$cons >0), ])
nc[3,1] = nrow(dat[which(dat$gdp< tp[3,1] & dat$gdp >0), ])
nc[4,1] = nrow(dat[which(dat$cons< tp[4,1] & dat$cons >0), ])

nc.pc[1,1] = round(nc[1,1]/n.gdp*100,0)
nc.pc[2,1] = round(nc[2,1]/n.cons*100,0)
nc.pc[3,1] = round(nc[3,1]/n.gdp*100,0)
nc.pc[4,1] = round(nc[4,1]/n.cons*100,0)

nc[1,2] = nrow(dat[which(dat$gdp< tp[1,2] & dat$gdp >0), ])
nc[2,2] = nrow(dat[which(dat$cons< tp[2,2] & dat$cons >0), ])
nc[3,2] = nrow(dat[which(dat$gdp< tp[3,2] & dat$gdp >0), ])
nc[4,2] = nrow(dat[which(dat$cons< tp[4,2] & dat$cons >0), ])

nc.pc[1,2] = round(nc[1,2]/n.gdp*100,0)
nc.pc[2,2] = round(nc[2,2]/n.cons*100,0)
nc.pc[3,2] = round(nc[3,2]/n.gdp*100,0)
nc.pc[4,2] = round(nc[4,2]/n.cons*100,0)

nc[1,3] = nrow(dat[which(dat$gdp< tp[1,3] & dat$gdp >0), ])
nc[2,3] = nrow(dat[which(dat$cons< tp[2,3] & dat$cons >0), ])
nc[3,3] = nrow(dat[which(dat$gdp< tp[3,3] & dat$gdp >0), ])
nc[4,3] = nrow(dat[which(dat$cons< tp[4,3] & dat$cons >0), ])

nc.pc[1,3] = round(nc[1,3]/n.gdp*100,0)
nc.pc[2,3] = round(nc[2,3]/n.cons*100,0)
nc.pc[3,3] = round(nc[3,3]/n.gdp*100,0)
nc.pc[4,3] = round(nc[4,3]/n.cons*100,0)

nc[1,4] = nrow(dat[which(dat$gdp< tp[1,4] & dat$gdp >0), ])
nc[2,4] = nrow(dat[which(dat$cons< tp[2,4] & dat$cons >0), ])
nc[3,4] = nrow(dat[which(dat$gdp< tp[3,4] & dat$gdp >0), ])
nc[4,4] = nrow(dat[which(dat$cons< tp[4,4] & dat$cons >0), ])

nc.pc[1,4] = round(nc[1,4]/n.gdp*100,0)
nc.pc[2,4] = round(nc[2,4]/n.cons*100,0)
nc.pc[3,4] = round(nc[3,4]/n.gdp*100,0)
nc.pc[4,4] = round(nc[4,4]/n.cons*100,0)

nc[1,5] = nrow(dat[which(dat$gdp< tp[1,5] & dat$gdp >0), ])
nc[2,5] = nrow(dat[which(dat$cons< tp[2,5] & dat$cons >0), ])
nc[3,5] = nrow(dat[which(dat$gdp< tp[3,5] & dat$gdp >0), ])
nc[4,5] = nrow(dat[which(dat$cons< tp[4,5] & dat$cons >0), ])

nc.pc[1,5] = round(nc[1,5]/n.gdp*100,0)
nc.pc[2,5] = round(nc[2,5]/n.cons*100,0)
nc.pc[3,5] = round(nc[3,5]/n.gdp*100,0)
nc.pc[4,5] = round(nc[4,5]/n.cons*100,0)

nc[1,6] = nrow(dat[which(dat$gdp< tp[1,6] & dat$gdp >0), ])
nc[2,6] = nrow(dat[which(dat$cons< tp[2,6] & dat$cons >0), ])
nc[3,6] = nrow(dat[which(dat$gdp< tp[3,6] & dat$gdp >0), ])
nc[4,6] = nrow(dat[which(dat$cons< tp[4,6] & dat$cons >0), ])

nc.pc[1,6] = round(nc[1,6]/n.gdp*100,0)
nc.pc[2,6] = round(nc[2,6]/n.cons*100,0)
nc.pc[3,6] = round(nc[3,6]/n.gdp*100,0)
nc.pc[4,6] = round(nc[4,6]/n.cons*100,0)

nc[1,7] = nrow(dat[which(dat$gdp< tp[1,7] & dat$gdp >0), ])
nc[2,7] = nrow(dat[which(dat$cons< tp[2,7] & dat$cons >0), ])
nc[3,7] = nrow(dat[which(dat$gdp< tp[3,7] & dat$gdp >0), ])
nc[4,7] = nrow(dat[which(dat$cons< tp[4,7] & dat$cons >0), ])

nc.pc[1,7] = round(nc[1,7]/n.gdp*100,0)
nc.pc[2,7] = round(nc[2,7]/n.cons*100,0)
nc.pc[3,7] = round(nc[3,7]/n.gdp*100,0)
nc.pc[4,7] = round(nc[4,7]/n.cons*100,0)

nc.pc.colmean = colMeans(nc.pc)
nc.pc.mean = mean(nc.pc.colmean)
nc.pc.panel.mean = mean(nc.pc.colmean[6:7])
nc.pc.nonpanel.mean = mean(nc.pc.colmean[-(6:7)])
@

\begin{table}
\caption{Estimated turning points}
\label{tab:tp}
\small
\centering
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Table & Estimator & dep. var & indep. var & tp (\$, PPP) & sign. & \# ctrs left of tp.\tabularnewline
\hline 
\hline
\ref{tab:fpi-regressions} & OLS & FPI infl. 07-08  & GDP & \Sexpr{round(tp[1,1],0)} & yes & \Sexpr{nc[1,1]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,1]} \%) \tabularnewline
\hline 
\ref{tab:fpi-regressions} & OLS & FPI infl. 07-08  & Cons & \Sexpr{round(tp[2,1],0)} & yes & \Sexpr{nc[2,1]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,1]} \%) \tabularnewline
\hline 
\ref{tab:fpi-regressions} & OLS & FPI infl. 10-11  & GDP & \Sexpr{round(tp[3,1],0)} & no & \Sexpr{nc[3,1]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,1]} \%) \tabularnewline
\hline 
\ref{tab:fpi-regressions} & OLS & FPI infl. 10-11  & Cons & \Sexpr{round(tp[4,1],0)} & yes & \Sexpr{nc[4,1]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,1]} \%) \tabularnewline
\hline 
\ref{tab:rfpi-regressions} & OLS & rFPI growth 07-08  & GDP & \Sexpr{round(tp[1,2],0)} & yes & \Sexpr{nc[1,2]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,2]} \%) \tabularnewline
\hline 
\ref{tab:rfpi-regressions} & OLS & rFPI growth 07-08  & Cons & \Sexpr{round(tp[2,2],0)} & yes & \Sexpr{nc[2,2]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,2]} \%) \tabularnewline
\hline 
\ref{tab:rfpi-regressions} & OLS & rFPI growth 10-11  & GDP & \Sexpr{round(tp[3,2],0)} & yes & \Sexpr{nc[3,2]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,2]} \%) \tabularnewline
\hline 
\ref{tab:rfpi-regressions} & OLS & rFPI growth 10-11  & Cons & \Sexpr{round(tp[4,2],0)} & yes & \Sexpr{nc[4,2]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,2]} \%) \tabularnewline
\hline 
\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & GDP & \Sexpr{round(tp[1,3],0)} & yes & \Sexpr{nc[1,3]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,3]} \%) \tabularnewline
\hline 
\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & Cons & \Sexpr{round(tp[2,3],0)} & yes & \Sexpr{nc[2,3]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,3]} \%) \tabularnewline
\hline 
\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & GDP & \Sexpr{round(tp[3,3],0)} & yes & \Sexpr{nc[3,3]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,3]} \%) \tabularnewline
\hline 
\ref{tab:int-regressions} & OLS & FPI infl. 07-08  & Cons & \Sexpr{round(tp[4,3],0)} & yes & \Sexpr{nc[4,3]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,3]} \%) \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & OLS & FPI LRM  & GDP & \Sexpr{round(tp[1,4],0)} & yes & \Sexpr{nc[1,4]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,4]} \%) \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & OLS & FPI LRM  & Cons & \Sexpr{round(tp[2,4],0)} & yes & \Sexpr{nc[2,4]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,4]} \%) \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & WLS & FPI LRM  & GDP & \Sexpr{round(tp[3,4],0)} & no & \Sexpr{nc[3,4]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,4]} \%) \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & WLS & FPI LRM  & Cons & \Sexpr{round(tp[4,4],0)} & yes & \Sexpr{nc[4,4]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,4]} \%) \tabularnewline
\hline
\ref{tab:rfpi-pt-regressions} & OLS & rFPI LRM  & GDP & \Sexpr{round(tp[1,5],0)} & yes & \Sexpr{nc[1,5]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,5]} \%) \tabularnewline
\hline 
\ref{tab:rfpi-pt-regressions} & OLS & rFPI LRM  & Cons & \Sexpr{round(tp[2,5],0)} & yes & \Sexpr{nc[2,5]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,5]} \%) \tabularnewline
\hline 
\ref{tab:rfpi-pt-regressions} & WLS & rFPI LRM  & GDP & \Sexpr{round(tp[3,5],0)} & no & \Sexpr{nc[3,5]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,5]} \%) \tabularnewline
\hline 
\ref{tab:rfpi-pt-regressions} & WLS & rFPI LRM  & Cons & \Sexpr{round(tp[4,5],0)} & no & \Sexpr{nc[4,5]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,5]} \%) \tabularnewline
\hline
\ref{tab:fe-panel-regressions} & FE & FPI infl.  & GDP & \Sexpr{round(tp[1,6],0)} & yes & \Sexpr{nc[1,6]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,6]} \%) \tabularnewline
\hline 
\ref{tab:fe-panel-regressions} & FE & FPI infl.  & Cons & \Sexpr{round(tp[2,6],0)} & yes & \Sexpr{nc[2,6]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,6]} \%) \tabularnewline
\hline 
\ref{tab:fe-panel-regressions} & FE & rFPI growth  & GDP & \Sexpr{round(tp[3,6],0)} & yes & \Sexpr{nc[3,6]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,6]} \%) \tabularnewline
\hline 
\ref{tab:fe-panel-regressions} & FE & rFPI growth  & Cons & \Sexpr{round(tp[4,6],0)} & yes & \Sexpr{nc[4,6]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,6]} \%) \tabularnewline
\hline
\ref{tab:fd-panel-regressions} & AH & FPI infl.  & GDP & \Sexpr{round(tp[1,7],0)} & no & \Sexpr{nc[1,7]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[1,7]} \%) \tabularnewline
\hline 
\ref{tab:fd-panel-regressions} & AH & FPI infl.  & Cons & \Sexpr{round(tp[2,7],0)} & no & \Sexpr{nc[2,7]}/\Sexpr{n.cons} (\Sexpr{nc.pc[2,7]} \%) \tabularnewline
\hline 
\ref{tab:fd-panel-regressions} & AH & rFPI growth  & GDP & \Sexpr{round(tp[3,7],0)} & no & \Sexpr{nc[3,7]}/\Sexpr{n.gdp} (\Sexpr{nc.pc[3,7]} \%) \tabularnewline
\hline 
\ref{tab:fd-panel-regressions} & AH & rFPI growth  & Cons & \Sexpr{round(tp[4,7],0)} & yes & \Sexpr{nc[4,7]}/\Sexpr{n.cons} (\Sexpr{nc.pc[4,7]} \%) \tabularnewline
\hline 
\end{tabular}
\end {table}