### 2016-03-02 data_prep.Rnw
Revised data preparation script. Used for 2. round version of paper

% !Rnw root = Master.Rnw

<<datasets, echo=F>>=
library(plyr)
library(dplyr)
library(zoo)
library(plm)
library(foreign)
library(reshape)
library(reshape2)
library(ggplot2)
library(data.table)
library(lattice)
library(readstata13)
library(countrycode)

### Data files used
# laborstaB9.csv
# laborstaBA.csv

#####################################################################

str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"

### The ILO data
temp = paste(str, "laborstaB9.csv", sep = "")
B9 = read.csv(temp, sep = ";")
temp = paste(str, "laborstaBA.csv", sep = "")
BA = read.csv(temp, sep = ";")

## Reshape to long format
B9$id = paste0(B9$CODE.COUNTRY, B9$CODE.AREA)
B9[B9$id == "LKLK1" & B9$YEAR == 2002, ]
B9 = B9[-2659, ] # coding mistake
B9l = reshape(B9, varying=c(9:21), direction="long", v.names="cpi", 
              idvar=c("id", "YEAR"),
              sep="", timevar = "month")
B9l = B9l[order(B9l$id, B9l$YEAR, B9l$month), ]
B9l$SRC_ID = NULL
B9l = rename(B9l, c("NOTES" = "cpi.notes"))

BA$id = paste0(BA$CODE.COUNTRY, BA$CODE.AREA)
BA[BA$id == "LKLK1" & BA$YEAR == 2002, ]
BA = BA[-2629, ] # coding mistake
BAl = reshape(BA, varying=c(9:21), direction="long", v.names="fpi", 
              idvar=c("id", "YEAR"),
              sep="", timevar = "month")
BAl = BAl[order(BAl$id, BAl$YEAR, BAl$month), ]
BAl$SRC_ID = NULL
BAl = rename(BAl, c("NOTES" = "fpi.notes"))

## Merge CPI and FPI datasets
ilo = merge(B9l, BAl, all=TRUE)
ilo = ilo[order(ilo$id, ilo$YEAR, ilo$month), ]
ilo = rename(ilo, c("YEAR" = "year", "COUNTRY"="country", 
                    "CODE.COUNTRY"="c.code", "CODE.AREA"="a.code"))

ilo = ilo[ilo$month != 13, ]
ilo$tt = as.yearmon(paste(ilo$year, ilo$month, sep = "-"))

rm(B9, BA, B9l, BAl)

## Identifying countries with multiple price series and select single series
ilo$id = factor(ilo$id)
str(ilo$id)
summary(tabulate(ilo$id))
table(ilo$id)
table(table(ilo$id))
ilo$c.code = factor(ilo$c.code)
str(ilo$c.code)
table(table(ilo$c.code))
tbl = table(ilo$c.code)
# subset of countries with > 1 price serie
dat = droplevels(ilo[ilo$c.code %in% names(tbl)[tbl > 180],,drop=FALSE])
table(dat$c.code)
length(table(dat$c.code))

# visual inspection of the series
c = "MM"
datsub = ilo[ilo$c.code == c, ]
dat.m = melt(datsub,measure.vars=c('cpi','fpi'))
ggplot(dat.m) +
 geom_line(aes(x=as.numeric(tt),y=value,group=id,color=id)) +
  facet_grid(~variable)

# drop/keep (ir)relevant series
ilo = ilo[ilo$id != "BRBR1", ]
ilo = ilo[ilo$c.code != "BS", ]
ilo = ilo[ilo$c.code != "CL", ]
ilo = ilo[ilo$c.code != "CV", ]
ilo = ilo[ilo$id != "ETET1", ]
ilo = ilo[!(ilo$id %in% c("ININ4", "ININ6", "ININ7")), ]
ilo = ilo[ilo$c.code != "KE", ]
ilo = ilo[ilo$c.code != "LB", ]
ilo = ilo[!(ilo$id %in% c("MGMG1", "MGMG2")), ]
ilo = ilo[ilo$c.code != "MN", ]
ilo = ilo[ilo$id != "MZMZ1", ]
ilo = ilo[ilo$id != "NINI1", ]
ilo = ilo[ilo$id != "OMOM1", ]
ilo = ilo[ilo$id != "PAPA1", ]
ilo = ilo[ilo$id != "RW", ]
ilo = ilo[ilo$id != "SASA2", ]
ilo = ilo[ilo$id != "SLSL1", ]
ilo = ilo[ilo$id != "SN", ]
ilo = ilo[ilo$id != "SYSY1", ]
ilo = ilo[ilo$c.code != "VE", ]
# Additional countries with multiple series
ilo = ilo[ilo$id != "ZMZM2", ]
ilo = ilo[ilo$c.code != "KN", ]
ilo = ilo[ilo$c.code != "SZ", ]
ilo = ilo[ilo$id != "MMMM1", ]

ilo$id = factor(ilo$id)
str(ilo$id)
ilo$c.code = factor(ilo$c.code)
str(ilo$c.code)
rm(dat, dat.m, datsub)
ilo$id = NULL
ilo = rename(ilo, c("c.code" = "id"))

## Selecting the gross sample of countries
x = 180 -(3*12)
x/12 # number of years required
dat = droplevels(ilo[ilo$id %in% names(tbl)[tbl >= x],,drop=FALSE])
dat$id = factor(dat$id)
str(dat$id)
# require comple series
dat = dat %>% dplyr::filter(ave(!is.na(cpi), id, FUN = all))
dat = dat %>% dplyr::filter(ave(!is.na(fpi), id, FUN = all))
dat$id = factor(dat$id)
str(dat$id)
ilo = dat
rm(dat)

## Identifying and correcting for technical breaks, coding errors etc.
DT = data.table(ilo)
DT[, temp1 := 0]
setkey(DT, id)
DT[DT[, .I[1], by = key(DT)]$V1, temp1 := 2] 
DT[DT[, .I[.N], by = key(DT)]$V1, temp1 := 1]
DF = data.frame(DT)
DF = within(DF, lag.cpi.notes <- c(NA, head(as.character(cpi.notes), -1)))
DF = within(DF, lag.cpi <- c(NA, head((cpi), -1)))

DF = within(DF, lag.fpi.notes <- c(NA, head(as.character(fpi.notes), -1)))
DF = within(DF, lag.fpi <- c(NA, head((fpi), -1)))

DF$cpi.brk = factor(ifelse(as.character(DF$cpi.notes)!=as.character(DF$lag.cpi.notes) & 
                             is.na(DF$cpi.notes) == FALSE &
                             is.na(DF$lag.cpi.notes) == FALSE &
                             DF$temp1 !=2 &
                             abs(DF$cpi/DF$lag.cpi-1) > 0.1, 
                           1,0))
DF$fpi.brk = factor(ifelse(as.character(DF$fpi.notes)!=as.character(DF$lag.fpi.notes) & 
                             is.na(DF$fpi.notes) == FALSE &
                             is.na(DF$lag.fpi.notes) == FALSE &
                             DF$temp1 !=2 &
                             abs(DF$fpi/DF$lag.fpi-1) > 0.1, 
                           1,0))
summary(DF$cpi.brk)
summary(DF$fpi.brk)

View(DF[which((DF$cpi.brk) == 1),])
View(DF[which((DF$fpi.brk) == 1),])

DF$adj.cpi = ilo$cpi
DF$adj.fpi = ilo$fpi

# fix coding errors
DF[which(DF$id == "MD" & DF$year == 2007 & DF$month == 1), "cpi"] = 1195.14931/10

# fix the structural breaks
for (c in levels(DF$id)) 
  {
  temp = filter(DF, id == c)
  temp$d.cpi = c(NA, diff(temp$cpi))
  temp$x = ifelse(temp$cpi.brk == 1, temp$d.cpi, 0)
  temp = within(temp, cs.cpi <- cumsum(x))
  temp$adj.cpi = temp$cpi-temp$cs.cpi

  temp$d.fpi = c(NA, diff(temp$fpi))
  temp$y = ifelse(temp$fpi.brk == 1, temp$d.fpi, 0)
  temp = within(temp, cs.fpi <- cumsum(y))
  temp$adj.fpi = temp$fpi-temp$cs.fpi

  DF[which(DF$id == c), c("adj.cpi", "adj.fpi")] = temp[, c("adj.cpi", "adj.fpi")]
  }

# visual inspection of the series with structural breaks
c = "AM"
datsub = DF[DF$id == c, ]
dat.m = melt(datsub,measure.vars=c('cpi','fpi'))
ggplot(dat.m) +
 geom_line(aes(x=as.numeric(tt),y=value,group=id,color=id)) +
  facet_grid(~variable)

dat.m = melt(datsub,measure.vars=c('adj.cpi','adj.fpi'))
ggplot(dat.m) +
 geom_line(aes(x=as.numeric(tt),y=value,group=id,color=id)) +
  facet_grid(~variable)

View(DF[which((DF$id) == c),])

## finishing touches
ilo = select(DF, country:tt, cpi.brk, fpi.brk, adj.cpi, adj.fpi)
rm(dat.m, datsub, DF, DT, temp)

ilo = mutate(ilo, rfpi = adj.fpi/adj.cpi*100)

# visual inspection of the series
xyplot(rfpi ~ tt | id,
       data = ilo[which(as.numeric(ilo$id)>=97 & as.numeric(ilo$id)<109), ],
       type = "l",
       col = c("black", "red"))

# standardising country names
ilo$iso3c = factor(countrycode(ilo$id, "iso2c", "iso3c", warn = TRUE))
ilo = ilo[ilo$id != "T1", ]
str(ilo$iso3c)
ilo$id = factor(ilo$id)
str(ilo$id)

#####################################################################

### Covariates
temp = paste(str, "icp_exp_2005.dta", sep = "")
icp_data = read.dta13(temp)

temp = paste(str, "wb_pov.dta", sep = "")
wb_pov = read.dta13(temp)

temp = paste(str, "fao_fpi.csv", sep = "")
fao_fpi = read.csv(temp, sep = ";")

temp = paste(str, "staples.Rda", sep = "")
load(temp)

temp = paste(str, "cdep.Rda", sep = "")
load(temp)

temp = paste(str, "infl_0708.dta", sep = "")
infl_0708 =  read.dta13(temp)

temp = paste(str, "infl_1011.dta", sep = "")
infl_1011 =  read.dta13(temp)

df = ilo_data[,c("id", "year", "month", "acpi", "afpi", "rafpi", "iso3c")]
df1 = merge(df, icp_data, by = "iso3c", all.x=TRUE)
df1 = merge(df1, wb_pov, by = "iso3c", all.x=TRUE)
df1 = merge(df1, staples, by = "iso3c", all.x=TRUE)
df1 = merge(df1, cdep, by = "iso3c", all.x=TRUE)

df1$tt = as.yearmon(paste(df1$year, df1$month, sep = "-"))
df1 = df1[order(df1$id, df1$tt),]
df2 = cbind(df1, fao_fpi)
data = plm.data(df2, index = c("id", "tt"))

quar_gdp = quantile(icp_data$GDP, c(.25, .5, .75), na.rm = T)
fao_0708 = (fao_fpi[42,2]/fao_fpi[25,2]-1)*100
fao_1011 = (fao_fpi[72,2]/fao_fpi[61,2]-1)*100

infl_0708 = rename(infl_0708, c(id="iso3c"))
data_0708 = merge(infl_0708, icp_data, by = "iso3c", all.x=TRUE)
data_0708 = merge(data_0708, wb_pov, by = "iso3c", all.x=TRUE)
data_0708 = merge(data_0708, staples, by = "iso3c", all.x=TRUE)
data_0708 = merge(data_0708, cdep, by = "iso3c", all.x=TRUE)
data_0708 = cbind(data_0708, fao_0708)

infl_1011 = rename(infl_1011, c(id="iso3c"))
data_1011 = merge(infl_1011, icp_data, by = "iso3c", all.x=TRUE)
data_1011 = merge(data_1011, wb_pov, by = "iso3c", all.x=TRUE)
data_1011 = merge(data_1011, staples, by = "iso3c", all.x=TRUE)
data_1011 = merge(data_1011, cdep, by = "iso3c", all.x=TRUE)
data_1011 = cbind(data_1011, fao_1011)

infl = merge(infl_0708, infl_1011, by = "iso3c")
infl_long = reshape(infl,varying=c("foodinfl78","foodinfl1011","cpiinfl78", "cpiinfl1011", 
                                  "rfpigr78", "rfpigr1011"), direction="long", 
                   idvar="iso3c", sep="")
infl_long1 = melt(infl_long, id = c("iso3c", "time"))
@