### 2016-03-02 figures.Rnw
Revised figures script. Used for 2. round rev. version of paper

% !Rnw root = Master.Rnw

\SweaveInput{data_prep.Rnw}

\section{Figures}
<<pt, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
a = 1.3
b = 1.2
par(mfrow=c(1,2))
curve(1-a^(-x), 0, 10, lty=1, xlab = "Income", ylab = "", ylim = c(0,1))
curve(b^(-x), add = T, col = "blue", lty=2)
title(main = "Shares", cex.main = 1)
legend("topright", c(expression(beta(y)), expression(gamma(y))),lty = c(2,1), col = c("blue", "black"))
curve((1-a^(-x))*b^(-x), 0, 10, xlab = "Income", ylab = "")
title(main = "Price transmission", cex.main = 1)
legend("topright", c(expression(theta(y))), lty = 1, col = "black")
@

<<gamma, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
library(ggplot2)
library("foreign")

## Import data into R.
str = "C:/Users/nqw235/Google Drev/Research/pov_price_transmission/"
temp = paste(str, "dist_trade.csv", sep = "")

dat = read.csv(temp, sep = ";")
dat$gamma = (dat$vop_covx+dat$vop_covm)/dat$vop_covt*100
ggplot(dat, aes(log(gdppcp00), gamma)) + 
  geom_point() + 
  geom_smooth() +
  theme_bw() +
  xlab("Log of GDP") +
  ylab(expression(paste("Estimated ", gamma))) +
  theme(panel.grid.major = element_line(colour = "grey"))
@

<<infl, echo=FALSE, include=FALSE, fig=TRUE, height=6, width=8>>=
library(gridExtra)

p1 = ggplot(aes(y = value, x = factor(time, labels = c("2007:1-2008:6", "2010:1-2010:12")), 
                fill = factor(variable, labels = c("FPI inflation", "CPI inflation", "Real FPI growth"))), data = infl_long1) + 
  geom_boxplot(outlier.shape = 1) +
  theme_bw() +
  scale_fill_grey(start = 0, end = 1) +
  theme(legend.position=c(.75,.775)) +
  labs(fill = "") +
  ylab("Annualized growth rates (%)") + 
  xlab("")

p2 = ggplot(aes(x=foodinfl78, y=foodinfl1011), data = infl) + 
  geom_point(shape=1) +
  geom_smooth(method=loess) +
  theme_bw() +
  ylab("FPI inflation 2010-11") + 
  xlab("FPI inflation 2007-08") +
  geom_text(aes(label=ifelse(foodinfl78>40,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
            size=3) +
  geom_text(aes(label=ifelse(foodinfl1011>20,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
            size=3)

p3 = ggplot(aes(x=cpiinfl78, y=cpiinfl1011), data = infl) +
  geom_point(shape=1) +
  geom_smooth(method=loess) +
  theme_bw() +
  ylab("CPI inflation 2010-11") + 
  xlab("CPI inflation 2007-08") +
  geom_text(aes(label=ifelse(cpiinfl78>30,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
            size=3) +
  geom_text(aes(label=ifelse(cpiinfl1011>15,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
            size=3)

p4 = ggplot(aes(x=rfpigr78, y=rfpigr1011), data = infl) + 
  geom_point(shape=1) +
  geom_smooth(method=loess) +
  theme_bw() +
  ylab("Real FPI growth 2010-11") + 
  xlab("Real FPI Growth 2007-08") +
  geom_text(aes(label=ifelse(rfpigr78>12,as.character(iso3c),'')),hjust=1.2,vjust=-1.2, 
            size=3) +
  geom_text(aes(label=ifelse(rfpigr1011>10,as.character(iso3c),'')),hjust=1.2,vjust=-1.2,
            size=3)

grid.arrange(p1, p2, p3, p4, ncol=2)
@

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{Master-pt.pdf}
\caption{The effect of income on price transmission}
\label{fig:pt}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{Master-gamma.pdf}
\caption{Estimates of $\gamma$. Source: \cite{Anderson2008}. Own calculations.}
\label{fig:gamma}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{Master-infl.pdf}
\caption{Domestic price changes 2007-8 and 2010-11. Source: ILO, own calculations.}
\label{fig:infl}
\end{center}
\end{figure}