### 2016-08-11 tables.Rnw
JAE sub. vers.

% !Rnw root = Master.Rnw

\section*{Tables}

<<results=tex,echo=F>>=
str = "C:/Users/nqw235/Google Drive/Research/pov_price_transmission/"
library(stargazer)
reg.list = readRDS(paste0(str,"fc_reg0708_m.rds"))
se.list = readRDS(paste0(str,"fc_reg0708_se.rds"))

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of domestic food price changes 2007:1-2008:6', 
          label = "tab:07-08-fc-regressions", 
          dep.var.labels=c("FPI inflation", "RFPI growth"),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          keep.stat = c("rsq", "n", "aic"),
          font.size = "small", 
          covariate.labels=c("$\\Delta ER$",
                             "$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
                  "percent levels, respectively. Robust standard errors in;", 
                  "parentheses samples exclude outliers defined as obser-", 
                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
                  "dependent variable. A description of the variables and", 
                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex,echo=F>>=
library(stargazer)
reg.list = readRDS(paste0(str,"fpi_pt_reg_m.rds"))
se.list = readRDS(paste0(str,"fpi_pt_reg_se.rds"))
bp.stat.list = readRDS(paste0(str,"fpi_bptest_stat.rds"))
bp.p.list = readRDS(paste0(str,"fpi_bptest_p.rds"))

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of long run FPI transmission rates', 
          label = "tab:fpi-pt-regressions", 
          dep.var.labels=c("FPI LRM (6 lags)"), 
          column.labels = c("OLS", "WLS"),
          column.separate = c(2, 2),
          no.space=TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]), 
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small", 
          covariate.labels=c("$GDP$",
                             "$GDP^{2}$", 
                             "$CCONS$", 
                             "$CIMP$", 
                             "$CEXP$",
                             "$LLDC$", 
                             "$STE$",
                             "$TO$", 
                             "$DTF$", 
                             "$LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
                  "percent levels, respectively. Robust standard errors in;", 
                  "parentheses samples exclude outliers defined as obser-", 
                  "vations $\\pm$ 3.5 standard deviations from the mean of the", 
                  "dependent variable. A description of the variables and", 
                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<results=tex, echo=FALSE>>=
library(plm)
reg.list = readRDS(paste0(str,"panel_reg_m.rds"))
se.list = readRDS(paste0(str,"panel_reg_se.rds"))

stargazer(reg.list[[1]], reg.list[[2]], reg.list[[3]], reg.list[[4]],
          title='Determinants of monthly FPI inflation and RFPI growth', 
          label = "tab:panel-regressions", 
          dep.var.labels=c("FPI inflation", "RFPI growth"),
          no.space = TRUE,
          se = list(se.list[[1]], se.list[[2]], se.list[[3]], se.list[[4]]),
          omit = c("adj.fpi", "rfpi"),
          omit.stat = c("adj.rsq","f", "ser"), 
          font.size = "small",
          covariate.labels=c("$\\pi_{t}^{i}$",
                             "$\\pi_{t-1}^{i}$",
                             "$\\Delta ER_{t}$",
                             "$\\Delta ER_{t-1}$",
                             "$\\pi_{t-1}^{i}\\times GDP$",
                             "$\\pi_{t-1}^{i}\\times GDP^2$", 
                             "$\\pi_{t-1}^{i}\\times CCONS$", 
                             "$\\pi_{t-1}^{i}\\times CIMP$", 
                             "$\\pi_{t-1}^{i}\\times CEXP$",
                             "$\\pi_{t-1}^{i}\\times LLDC$", 
                             "$\\pi_{t-1}^{i}\\times STE$", 
                             "$\\pi_{t-1}^{i}\\times TO$", 
                             "$\\pi_{t-1}^{i}\\times DTF$", 
                             "$\\pi_{t-1}^{i}\\times LPI$"),
          notes=c("*, ** and *** signify significance at the 10, 5 and 1",
                  "percent levels, respectively. Robust standard errors", 
                  "in parentheses. A description of the variables and", 
                  "their sources can be found in table \\ref{tab:cs_data} and section \\ref{sec:analysis}."),
          notes.append=FALSE,
          notes.align="l")
@

<<echo=FALSE>>=
load(paste0(str,"infl_ctr_dat_0708.Rda"))
dat = ctr.dat.0708
r1 = readRDS(paste0(str,"fc_reg0708_m.rds"))
r2 = readRDS(paste0(str,"fpi_pt_reg_m.rds"))
r3 = readRDS(paste0(str,"panel_reg_m.rds"))

tp = matrix(NA, nrow=9, ncol=1)
tp[1,1] = -coef(r1[[1]])["log(gdp.const.int)"]/(2*coef(r1[[1]])["I(log(gdp.const.int)^2)"])
tp[2,1] = -coef(r1[[2]])["log(gdp.const.int)"]/(2*coef(r1[[2]])["I(log(gdp.const.int)^2)"])
tp[3,1] = -coef(r1[[3]])["log(gdp.const.int)"]/(2*coef(r1[[3]])["I(log(gdp.const.int)^2)"])
tp[4,1] = -coef(r1[[4]])["log(gdp.const.int)"]/(2*coef(r1[[4]])["I(log(gdp.const.int)^2)"])
tp[5,1] = -coef(r2[[1]])["log(gdp.const.int)"]/(2*coef(r2[[1]])["I(log(gdp.const.int)^2)"])
tp[6,1] = -coef(r2[[2]])["log(gdp.const.int)"]/(2*coef(r2[[2]])["I(log(gdp.const.int)^2)"])
tp[7,1] = -coef(r2[[3]])["log(gdp.const.int)"]/(2*coef(r2[[3]])["I(log(gdp.const.int)^2)"])
tp[8,1] = -coef(r2[[4]])["log(gdp.const.int)"]/(2*coef(r2[[4]])["I(log(gdp.const.int)^2)"])
tp[9,1] = -coef(r3[[1]])["I(lag(diff(log(fao.fpi)), 1) * log(gdp.const.int))"]/
  (2*coef(r3[[1]])["I(lag(diff(log(fao.fpi)), 1) * I(log(gdp.const.int)^2))"])

tp = exp(tp)
n.gdp = length(which(!is.na(dat$gdp.const.int)))

nc = matrix(NA, nrow=9, ncol=1)
nc[1,1] = nrow(dat[which(dat$gdp.const.int< tp[1,1] & !is.na(dat$gdp.const.int)), ])
nc[2,1] = nrow(dat[which(dat$gdp.const.int< tp[2,1] & !is.na(dat$gdp.const.int)), ])
nc[3,1] = nrow(dat[which(dat$gdp.const.int< tp[3,1] & !is.na(dat$gdp.const.int)), ])
nc[4,1] = nrow(dat[which(dat$gdp.const.int< tp[4,1] & !is.na(dat$gdp.const.int)), ])
nc[5,1] = nrow(dat[which(dat$gdp.const.int< tp[5,1] & !is.na(dat$gdp.const.int)), ])
nc[6,1] = nrow(dat[which(dat$gdp.const.int< tp[6,1] & !is.na(dat$gdp.const.int)), ])
nc[7,1] = nrow(dat[which(dat$gdp.const.int< tp[7,1] & !is.na(dat$gdp.const.int)), ])
nc[8,1] = nrow(dat[which(dat$gdp.const.int< tp[8,1] & !is.na(dat$gdp.const.int)), ])
nc[9,1] = nrow(dat[which(dat$gdp.const.int< tp[9,1] & !is.na(dat$gdp.const.int)), ])

nc.pc.mean =  matrix(NA, nrow=3, ncol=1)
nc.pc.mean[1] = mean(nc[1:9])/n.gdp
nc.pc.mean[2] = mean(nc[9])/n.gdp
nc.pc.mean[3] = mean(nc[1:8])/n.gdp
nc.pc.mean = nc.pc.mean*100

tp.mean =  matrix(NA, nrow=3, ncol=1)
tp.mean[1] = mean(tp[1:9])
tp.mean[2] = mean(tp[9])
tp.mean[3] = mean(tp[1:8])
save(nc.pc.mean, file = "tp_ctr_mean.Rda")
save(tp.mean, file = "tp_ctr_mean1.Rda")
@

\begin{table}
\centering
\begin{threeparttable}
\caption{Estimated turning points (TPs)}
\label{tab:tp}
\small
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline 
Table & Column & Estimator & Dependent variable & TP (\$, PPP) & Left of TP\tabularnewline
\hline 
\hline
\ref{tab:07-08-fc-regressions} & (1) & OLS & FPI infl. 07-08  &  \Sexpr{round(tp[1,1],0)} & \Sexpr{nc[1,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:07-08-fc-regressions} & (2) & OLS & FPI infl. 07-08  &  \Sexpr{round(tp[2,1],0)} & \Sexpr{nc[2,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:07-08-fc-regressions} & (3) & OLS & RFPI growth 07-08  &  \Sexpr{round(tp[3,1],0)} & \Sexpr{nc[3,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:07-08-fc-regressions} & (4) & OLS & RFPI growth 07-08  &  \Sexpr{round(tp[4,1],0)} & \Sexpr{nc[4,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & (1) & OLS & FPI LRM  &  \Sexpr{round(tp[5,1],0)} & \Sexpr{nc[5,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & (2) & WLS & FPI LRM  &  \Sexpr{round(tp[6,1],0)} & \Sexpr{nc[6,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & (3) & OLS & FPI LRM  &  \Sexpr{round(tp[7,1],0)} & \Sexpr{nc[7,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:fpi-pt-regressions} & (4) & WLS & FPI LRM  &  \Sexpr{round(tp[8,1],0)} & \Sexpr{nc[8,1]}/\Sexpr{n.gdp} \tabularnewline
\hline 
\ref{tab:panel-regressions} & (1) & FE & $\Delta \log(FPI)_{t}$  &  \Sexpr{round(tp[9,1],0)} & \Sexpr{nc[9,1]}/\Sexpr{n.gdp} \tabularnewline
\hline
\end{tabular}
\begin{tablenotes}
\item \emph{Note:} The numbers in the fourth column are the estimated turning points based on coefficients from the regression specified in the first three columns. I.e. the per capita income level corresponding to the highest level of predicted domestic food inflation or price transmission. The final column calculates how many countries that have per capita income less than the level corresponding to the estimated  turning point, relative to the total number of countries in the sample with data on income. 
\end{tablenotes}
\end{threeparttable}
\end{table}